
### Activity Monitoring Analysis

#### Read Data
```{r}
   setwd("~/Data Science Specialization/Reproducible Research/ProgrammingAssignment1")
   activity_data <- read.csv("activity.csv",sep=",")
```

### 1) What is mean total number of steps taken per day?

```{r} 
 
  totsteps <- aggregate(steps ~ date,activity_data,sum)

```
### Show two copies of plot indicating mean and median respectively 

```{r} 
 par(mfrow=c(2,1))

 hist(totsteps[,2], xlab="total steps taken per day", ylab="Frequency", breaks = 20, freq = TRUE, main="Total Steps per Day (Frequency)")
 abline(v=mean(totsteps[,2],na.rm=FALSE),col=4,lty=1,par(lwd=2))
 text( mean(totsteps[,2],na.rm=FALSE) + 1700,9,paste("---mean(",sprintf("%.2f",mean(totsteps[,2],na.rm=FALSE)),")"),col=4)

 hist(totsteps[,2], xlab="total steps taken per day", ylab="Frequency", breaks = 20, freq = TRUE, main="Total Steps per Day (Frequency)")
 abline(v=median(totsteps[,2],na.rm=FALSE),col=2,lty=1,par(lwd=2))
 text( median(totsteps[,2],na.rm=FALSE) + 2000,9,paste("---median(",median(totsteps[,2],na.rm=FALSE),")"),col=2)

```

### 2) What is the average Daily Activity Pattern?

```{r}
activity_data <- read.csv("activity.csv",sep=",")
 
totStepsPerInterval <- aggregate(steps ~ interval,activity_data,mean)

maxInterval <- subset(totStepsPerInterval,steps == max(steps))


```

```{r}

par(mfrow=c(1,1))

plot(totStepsPerInterval$interval, totStepsPerInterval$steps,type="s",lab=c(24,5,3))
abline(v=maxInterval$interval,col=3,lty=1,par(lwd=2))
text( maxInterval$interval + 150,200,paste("---maximum(",maxInterval$interval,")"),col=3)

```

### 3) What is the average Daily Activity Pattern?
### (3a How many missing values are there in the dataset ?)
```{r}
activity_data_with_NA <- activity_data[!complete.cases(activity_data),]
nrow(activity_data_with_NA) 

```
### 3b) Imputing missing values and graphing total steps taken showing mean/median aligned.

```{r}


 totStepsPerInterval <- aggregate(steps ~ interval,activity_data,mean)

 merged_activity_data <- merge(activity_data,totStepsPerInterval,by.x="interval",by.y="interval")
 merged_activity_data$steps.x[is.na(merged_activity_data$steps.x)] <- merged_activity_data$steps.y[is.na(merged_activity_data$steps.x)]
 revised_activity_data <- merged_activity_data[,1:3]

  totsteps <- aggregate(steps.x ~ date,revised_activity_data,sum)

 par(mfrow=c(2,1))

  hist(totsteps[,2], xlab="total steps taken per day", ylab="Frequency", breaks = 20, freq = TRUE, main="Total Steps per Day (Frequency)")
  abline(v=mean(totsteps[,2],na.rm=FALSE),col=4,lty=1,par(lwd=2)) 
  text( mean(totsteps[,2],na.rm=FALSE) + 1800,9,paste("---mean(",sprintf("%.2f",mean(totsteps[,2],na.rm=FALSE)),")"),col=4)

  hist(totsteps[,2], xlab="total steps taken per day", ylab="Frequency", breaks = 20, freq = TRUE, main="Total Steps per Day (Frequency)")
  abline(v=10765,col=2,lty=1,par(lwd=2))
  text( median(totsteps[,2],na.rm=FALSE) + 2100,9,paste("---median(",sprintf("%.2f",median(totsteps[,2],na.rm=FALSE)),")"),col=2)
```

